---
author: Goran S. Milovanovic"
date: "November 5, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# 'Introduction to R for Data Science'
## 'Session 01: Introduction to R from RStudio IDE'
### Organized by: [Data Science Serbia](http//:www.datascience.rs) and [Startit](http://en.startit.rs)
***

## Lecturers

#### [Goran S. Milovanović](http://www.exactness.net), Phd, Data Science Mentor at [Springboard](https://www.springboard.com/workshops/data-science), [Data Science Serbia](http://www.datascience.rs)
#### [ing Branko Kovač](https://rs.linkedin.com/in/kovacbranko), Data Scientist @Tradecore, Data Science Mentor at [Springboard](https://www.springboard.com/workshops/data-science), [Data Science Serbia](http://www.datascience.rs)

***
```{r echo=T}
# clear workspace
rm(list=ls())
# try this out:
ls()
```

```{r echo=T}
# R as a calculator
2+2
2/3-1+5
a <- 1/3
b = 7
a + b
```

```{r echo=T}
# what is my working directory?
getwd()
```

```{r echo=T}
# save R object as .rds file
saveRDS(a, "a.rds")
a <- readRDS("a.rds")
```

```{r echo=T}
# save several R objects as .RData file
save(list = c("a", "b"), file = "obj.RData")

# remove a, b
rm(a,b)
getwd()
# load .RDatafile to R
load(file = "obj.RData")
```

```{r echo=T}
# list of loaded objects
ls()

# remove all objects
rm(list=ls())
```

***
### How to install packages in RStudio

After you have successfully installed both **R** and **RStudio**, you are ready to start programming in R.
R has a plenty of built-in functions, but most of the time additional functions are needed. The true power of this programming language lies in the support from approximately 10,000 additional packages that can be installed to widen its functionality.

To install the packages that we are going to use during our course, please follow these steps:

1. Open RStudio from Start Menu (Windows users) or Applications folder (OS X users); Linux users: depends what desktop environment you use.

2. Assume that we want to install the package called: [dplyr](https://cran.r-project.org/web/packages/dplyr/index.html) (it is a well-known R package, indeed)

3. There are two ways to do it (actually, there are more than two ways, but let’s stick with the following two):

+ a. In the RStudio R console (the lower left quadrant of your RStudio workspace), type: *install.packages(‘dplyr’)*
+ b. From the *Tools* menu in RStudio, select *Instal Packages*, then start typing the package name and select it from the list

4. Most of the CRAN packages - those found in the standard, official R repository, called CRAN (Comprehensive R Archive Network) can be installed by following the standard procedure (described in Step 3). However, besides this basic package installation procedure, R packages from various sources can be installed in other ways as well. During the workshop we will demonstrate how to install R packages from the [GitHub](https://github.com/) repository (which is not R specific) using the [devtools](https://github.com/hadley/devtools) package from R, and how to install packages from the [Bioconductor](https://www.bioconductor.org/) repository.

5. *Note*: sometimes, most often in the R code comments, we use enclose the names of R packages in braces, like: {dplyr}, in order to visually mark the mention of an R package clearly.

6. *Note*: during the installation of some packages, you may notice how R starts installing additional packages - some packages that you have not selected for installation at all! The reason for this is the following: *some R packages rely on R functions developed in some other packages*. When an R package uses functions from another package, the later is called its **dependency**. R will automatically try to install all the dependencies of the package that you have selected for installation, so don’t be surprised if installing one package actually triggers and installation of a dozen of other packages at your system. It’s perfectly normal.

We will use many different R packages during the workshop. Here is a list of the packages that we will be needing; please take some time to practice package installation. To have the following packages pre-installed on the onset of our Workshop would really save us some time: [stringr](https://cran.r-project.org/web/packages/stringr/index.html), [stringi](https://cran.r-project.org/web/packages/stringi/index.html), [dplyr](https://cran.r-project.org/web/packages/dplyr/index.html), [tidyr](https://cran.r-project.org/web/packages/tidyr/index.html), [hflights](https://cran.rstudio.com/web/packages/hflights/index.html), [ggplot2](https://cran.r-project.org/web/packages/ggplot2/index.html), [corrplot](https://cran.r-project.org/web/packages/corrplot/index.html), [ppcor](https://cran.r-project.org/web/packages/ppcor/index.html), [Hmisc](https://cran.r-project.org/web/packages/Hmisc/index.html), [car](https://cran.r-project.org/web/packages/car/index.html), [QuantPsyc](https://cran.r-project.org/web/packages/QuantPsyc/index.html), [broom](https://cran.r-project.org/web/packages/broom/index.html), [lattice](https://cran.r-project.org/web/packages/lattice/index.html).

**REMINDER**: to install a package from the RStudio IDE, type in the console e.g.:

install.packages("dplyr")

```{r echo=T, results='hide'}
# installing packages:
# install.packages("dplyr") # replace "dplyr" with other package name
installed.packages() # list of installed packages
```

```{r echo=T}
# The shape of things to come:
numbers <- c(1,2,3,4,5)
words <- c("data", "science", "r")
boolean <- c(TRUE, FALSE)
data("mtcars")
```

***
#### [Data Science Serbia](http://www.datascience.rs)
